---
title: "Reproducible Research: Peer Assessment 1"
output:
html_document:
keep_md: true
---

### Loading and preprocessing the data
```{r Loading,   results="hide", warning=FALSE}
library(chron)
library(lattice)
setwd("C:/Users/Farhad/Documents/University/Coursera/Data Science  John Hopkins/4 - Reproducible Research/Homeworks")

Sys.time()
fURL="http://d396qusza40orc.cloudfront.net/repdata/data/activity.zip"
f<-read.csv(fURL,header=FALSE)
temp<-tempfile()
download.file(fURL,temp)
theData<-read.csv(unz(temp,"activity.csv"))
unlink(temp)

```

```{r some exploration,  echo=FALSE,  results="hide", warning=FALSE}
names(theData)
nrow(theData)
theData[1:9,]
unique(theData[,2])

```


```{r clean data, echo=FALSE, results="hide", warning=FALSE}
data<-theData
#data<-na.omit(theData[,c(2,3)])

data[,2]<-as.Date(data[,2])
data$day<-as.numeric(format(data[,2],"%d"))
data$month<-format(data[2],"%b")
dev.off()
windows()
```


###Steps taken each day

```{r Plot histogram, echo=FALSE, warning=FALSE}

octData<-subset(data,data$month=="Oct")
novData<-subset(data,data$month=="Nov")

#Plots
plotDataOct <- tapply(octData$steps, octData$day,  FUN = sum)
barplot(plotDataOct,col="red",xaxt="n",yaxt="n",ylab="",xlab="",space=1)
axis(side = 1, at = c(1:31), line=1, tick = FALSE, labels = FALSE,xlim=31,ylim=(0:16000))
#abline(lsfit(1:31, plotDataOct),col="red")
par(new=TRUE)
plotDataNov <- tapply(novData$steps, novData$day,  FUN = sum)
barplot(plotDataNov,col="blue",ylab= "Steps",pch=16,xlab="Days of the month",main="Sum of steps per day for October and November",space=2)
axis(side=1,at=c(0,100))
#abline(lsfit(1:30, plotDataNov),col="blue")
legend("bottomright",pch=c(1,16),col=c("red" ," blue"),legend=c("October","November"),cex=1,bg="white")
```



####Mean and Median of steps per day per month

```{r Mean & Meadian, echo=FALSE, results="hide", warning=FALSE}
#Remove zero steps values
NZoctData<-octData[octData$steps>0,]
NZnovData<-novData[novData$steps>0,]

NZoctData<-na.omit(NZoctData)
NZnovData<-na.omit(NZnovData)

#Get Oct values
MeanOct <- tapply(NZoctData$steps, NZoctData$day,  FUN = mean)
MedianOct <- tapply(NZoctData$steps, NZoctData$day,  FUN = median)
#Get Nov values
MeanNov <- tapply(NZnovData$steps, NZnovData$day,  FUN = mean)
MedianNov <- tapply(NZnovData$steps, NZnovData$day,  FUN = median)


OctSummary<-cbind(MeanOct,MedianOct)
NovSummary<-cbind(MeanNov,MedianNov)
```

```{r show mean and median, echo=FALSE,  warning=FALSE}
OctSummary
NovSummary



``` 



###Daily activity pattern (5 min intervals)   

```{r prepare data for plot, echo=FALSE, results="hide", warning=FALSE}
onData<-subset(data,data$month =="Oct" | data$month=="Nov")

agData<-aggregate(steps ~ interval, onData, FUN=sum)

```

```{r plot time interval, echo=FALSE,  warning=FALSE}
with(agData , plot(interval,steps,type="ln"))

```


####The maximum daily activity pattern   
```{r max steps, echo=FALSE, results="hide", warning=FALSE}
m<-max(agData$steps)
maxInt<-agData[agData$steps==m,1]
lmInt<-nchar(maxInt)
intM<-substr(maxInt,1,lmInt-2)
intS<-substr(maxInt,lmInt-1,lmInt)
tInt<- paste(intM , intS,"00",sep=":")
timeInt<-chron(times=tInt)

```
The maximum daily activity is `r maxInt` at `r timeInt`  


###Addressing missing Values    
`r nrow(theData) - nrow(na.omit(theData))` rows have "na" values.


```{r replace missing, echo=FALSE, results="hide", warning=FALSE}

ndata<-theData

#set all na values to 0
ndata$steps[is.na(ndata$steps)]<-0

#get an array of dates   
ad<-unique(theData$date)

#loop through dates
for (i in 1:length(ad)){
        #get the mean of all steps for a date
        mx<- mean(ndata[ndata$date==ad[i] & ndata$steps>0 ,1])
        #
        if (is.na(mx)) {
                mx<-mean(theData$steps,na.rm=TRUE)

        }
        ndata[ndata$date==as.character(ad[i]) & ndata$steps==0,1]<-  mx
        

}

```


```{r assign a mean to dates with mean equal to na, echo=FALSE, results="hide", warning=FALSE }
ndata[,2]<-as.Date(ndata[,2])
ndata$day<-as.numeric(format(ndata[,2],"%d"))
```
####Plot of 'Step' mean values with no missing value  
```{rplot mean , echo=FALSE,  warning=FALSE}

#Plots
plotData <- tapply(ndata$steps, ndata$day,  FUN = mean)
barplot(plotData,col="red",,ylab= "Steps",pch=16,xlab="Days of the month",main="Mean Steps per day")
axis(side = 1, at = c(1:31), line=1, tick = FALSE, labels = FALSE,xlim=31,ylim=(0:16000))

```

####Plot of 'Step' mean values with  no missing value in 5 min intervale   
```{r median , echo=FALSE,  warning=FALSE}
plotData <- tapply(ndata$steps, ndata$day,  FUN = median)
barplot(plotData,col="red",,ylab= "Steps",pch=16,xlab="Days of the month",main="Median Steps per day")
axis(side = 1, at = c(1:31), line=1, tick = FALSE, labels = FALSE,xlim=31,ylim=(0:16000))
```


###Plot of 'Step' mean values for weekdays in 5 min intervale

```{r create weekdays factor, echo=FALSE,  warning=FALSE}

#data$dayOfWeek<-weekdays(as.Date(data$date))
fDays<-c("Monday","Tuesday","Wednesday","Thursday","Friday")
fEnd<-c("Saturday","Sunday")


ndata$weekdays[weekdays(as.Date(ndata$date)) %in% fDays]<-"weekday"
ndata$weekdays[!(weekdays(as.Date(ndata$date)) %in% fDays)]<-"weekend"

dayData<-ndata[ndata$weekdays=="weekday",]
endData<-ndata[ndata$weekdays=="weekend",]


#Plots weekdays
par(mfrow=c(1,2))
plotData <- tapply(dayData$steps, dayData$day,  FUN = mean)
barplot(plotData,col="red",,ylab= "Steps",pch=16,xlab="Time interval",main="Mean Steps per Weekdays")
axis(side = 1, at = c(1:31), line=1, tick = FALSE, labels = FALSE,xlim=31,ylim=(0:16000))
#plot weekends
plotData <- tapply(endData$steps, endData$day,  FUN = mean)
barplot(plotData,col="red",,ylab= "Steps",pch=16,xlab="Time intervale",main="Mean Steps per weekends")
axis(side = 1, at = c(1:31), line=1, tick = FALSE, labels = FALSE,xlim=31,ylim=(0:16000))


```